---
title: "Analysis of Citibike Performance of eBikes vs Other Modes"
output: html_notebook
---

First, load all necessary packages.
```{r}
library("dplyr")      # Load dplyr package
library("plyr")       # Load plyr package
library("readr")      # Load readr package
library("purrr")      # Load purrr package
library("data.table") # Load data.table package for fread
```

Next, merge all .csv files into one dataframe.
```{r}
# From https://statisticsglobe.com/merge-csv-files-in-r

# Identify all CSV files
# CB_data <- list.files(path = "CitiBike_data/Raw/New_Format",  
#                        pattern = "*.csv", full.names = TRUE) %>% 
#   lapply(read_csv) %>%    # Store all files in list
#   bind_rows               # Combine data sets into one data set 
# head(CB_data)            # Print data to RStudio console

# Identify all CSV files
CB_data <- list.files(path = "CitiBike_data/Raw/New_Format",
                       pattern = "*.csv", full.names = TRUE) %>%
  lapply(read_csv) %>%            # Store all files in list
  reduce(full_join, by = "id")    # Full-join data sets into one data set
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

