---
title: "Analysis of Citibike Performance of eBikes vs Other Modes"
output: html_notebook
---

First, load all necessary packages.
```{r}
library(tidyverse)
library("dplyr")                  # Load dplyr package
# library("plyr")                 # Load plyr package
# library("readr")                # Load readr package
# library("purrr")                # Load purrr package
# install.packages("data.table")  # Install data.table package
# library("data.table")           # Load data.table package for fread
# install.packages("geosphere")   # Install geosphere package
#library("geosphere")             # Load geosphere package for distance calculation

# This part is to install the most updated package for "revgeo" directly from the developer
# Install the devtools package
# install.packages('devtools') 
# Detach the package already installed (if applicable); otherwise installing from github won't work
# detach("package:revgeo", unload = TRUE) 
# library(devtools)
# install_github('mhudecheck/revgeo')
# library("revgeo")
library(ggplot2)

# Prepare a color palette. Here with R color brewer:
library(RColorBrewer)
myPalette <- brewer.pal(5, "Set2") 
```
Next, create a pie chart of total Citibike rides (classic vs eBike)
```{r}
tot_rides_classic = as.numeric(table(CB_Data$rideable_type)[1])
tot_rides_electric = as.numeric(table(CB_Data$rideable_type)[2])

jpeg(filename="Visualizations/classic_vs_electric.jpg",width=1400,height=1200)
pie(c(tot_rides_classic, tot_rides_electric), 
    labels = c("Classic","Electric"), border="white", 
    col=c("blue","yellow"),
    cex=5)
dev.off()
```
Now, track the usage of classic vs eBikes over the course of the year
```{r}
# https://datavizpyr.com/bump-plot-with-ggbump/

time_series = c("May 2021", "June 2021", "July 2021", "August 2021",
                "September 2021", "October 2021,", "November 2021",
                "December 2021", "January 2022", "February 2022",
                "March 2022", "April 2022")

series_classic = c(
nrow(subset(CB_Data, rideable_type == "Classic Bike" & year == "2021" & month == "5", year)),
nrow(subset(CB_Data, rideable_type == "Classic Bike" & year == "2021" & month == "6", year)),
nrow(subset(CB_Data, rideable_type == "Classic Bike" & year == "2021" & month == "7", year)),
nrow(subset(CB_Data, rideable_type == "Classic Bike" & year == "2021" & month == "8", year)),
nrow(subset(CB_Data, rideable_type == "Classic Bike" & year == "2021" & month == "9", year)),
nrow(subset(CB_Data, rideable_type == "Classic Bike" & year == "2021" & month == "10", year)),
nrow(subset(CB_Data, rideable_type == "Classic Bike" & year == "2021" & month == "11", year)),
nrow(subset(CB_Data, rideable_type == "Classic Bike" & year == "2021" & month == "12", year)),
nrow(subset(CB_Data, rideable_type == "Classic Bike" & year == "2022" & month == "1", year)),
nrow(subset(CB_Data, rideable_type == "Classic Bike" & year == "2022" & month == "2", year)),
nrow(subset(CB_Data, rideable_type == "Classic Bike" & year == "2021" & month == "3", year)),
nrow(subset(CB_Data, rideable_type == "Classic Bike" & year == "2021" & month == "4", year))
)

series_electric = c(
nrow(subset(CB_Data, rideable_type == "Electric Bike" & year == "2021" & month == "5", year)),
nrow(subset(CB_Data, rideable_type == "Electric Bike" & year == "2021" & month == "6", year)),
nrow(subset(CB_Data, rideable_type == "Electric Bike" & year == "2021" & month == "7", year)),
nrow(subset(CB_Data, rideable_type == "Electric Bike" & year == "2021" & month == "8", year)),
nrow(subset(CB_Data, rideable_type == "Electric Bike" & year == "2021" & month == "9", year)),
nrow(subset(CB_Data, rideable_type == "Electric Bike" & year == "2021" & month == "10", year)),
nrow(subset(CB_Data, rideable_type == "Electric Bike" & year == "2021" & month == "11", year)),
nrow(subset(CB_Data, rideable_type == "Electric Bike" & year == "2021" & month == "12", year)),
nrow(subset(CB_Data, rideable_type == "Electric Bike" & year == "2022" & month == "1", year)),
nrow(subset(CB_Data, rideable_type == "Electric Bike" & year == "2022" & month == "2", year)),
nrow(subset(CB_Data, rideable_type == "Electric Bike" & year == "2021" & month == "3", year)),
nrow(subset(CB_Data, rideable_type == "Electric Bike" & year == "2021" & month == "4", year))
)

CB_usage_trend <- tibble(Time = time_series,
                        Classic = series_classic,
                        Electric = series_electric)

CB_usage_long <- CB_usage_trend %>%
  pivot_longer(-Time, names_to = "Type", values_to="Rides" )
 
CB_usage_long %>%
  ggplot(aes(Time, Rides, color = Type)) +
  geom_point(size = 5) +
  geom_bump() 
ggsave("Visualizations/bump_plot_ggbump.png", width=8, height=6)
```
Now, calculate actual duration of ride (in minutes) as well as year and month.
```{r}
setDT(CB_Data)[, year := format(as.Date(started_at), "%Y") ]
setDT(CB_Data)[, month := format(as.Date(started_at), "%m") ]
CB_Data$duration = as.numeric(
                              difftime(
                                CB_Data$ended_at, 
                                CB_Data$started_at, 
                                units ="mins"
                                )
                              )
head(CB_Data)
```
Figure out the distances being traveled. Since calculating actual roadmap distance through Google Maps or Bing require use of APIs that cost money and place limitations on daily computations, that process shall be forgone in favor of "as the crow flies", which will provide a rough estimate of travel distance in order to figure out speed.

Convert meters to miles.
```{r}
# https://stackoverflow.com/questions/49532911/calculate-distance-longitude-latitude-of-multiple-in-dataframe-r
CB_Data$distance <- 0.000621371 * distHaversine(
                              cbind(CB_Data$start_lng, CB_Data$start_lat), 
                              cbind(CB_Data$end_lng, CB_Data$end_lat) 
                              )
head(CB_Data)
```
Find the zip codes associated with the coordinate data. This will help with ascertaining neighborhood and borough locations for simplicity.
```{r}
CB_Data$start_zip <- revgeo(CB_Data$start_lng, CB_Data$start_lat,
                            provider = NULL, API = NULL, output = "hash",
                            item = "zip")
CB_Data$end_zip <- revgeo(CB_Data$end_lng, CB_Data$end_lat,
                            provider = NULL, API = NULL, output = "hash",
                            item = "zip")
# start_zip <- revgeo(-73.95772, 40.77863,
#                             provider = NULL, API = NULL, output = "hash",
#                             item = "zip")
# end_zip <- revgeo(-73.95578, 40.77280,
#                             provider = NULL, API = NULL, output = "hash",
#                             item = "zip")
# start_zip
# end_zip

head(CB_Data)
```

