---
title: "Analysis of Citibike Performance of eBikes vs Other Modes"
output: html_notebook
---

First, load all necessary packages.
```{r}
# library("dplyr")                # Load dplyr package
# library("plyr")                 # Load plyr package
# library("readr")                # Load readr package
# library("purrr")                # Load purrr package
# install.packages("data.table")  # Install data.table package
library("data.table")             # Load data.table package for fread
# install.packages("geosphere")   # Install geosphere package
library("geosphere")              # Load geosphere package for distance calculation

# This part is to install the most updated package for "revgeo" directly from the developer
# Install the devtools package
# install.packages('devtools') 
# Detach the package already installed (if applicable); otherwise installing from github won't work
# detach("package:revgeo", unload = TRUE) 
# library(devtools)
# install_github('mhudecheck/revgeo')
library("revgeo")
```
Next, read the .csv files that meet the latest format for CitiBike rider data.
```{r}
CB_Data <- fread("CitiBike_data/202102-202204-citibike-trip-data.csv")
```
Then, make sure all values for docked_bike and classic_bike are the same.
```{r}
CB_Data$rideable_type[CB_Data$rideable_type == "docked_bike"] <- "classic_bike"
# CB_Data <- replace(CB_Data$rideable_type, 
#                    CB_Data$rideable_type == "classic_bike",
#                    "Classic Bike")
# CB_Data <- replace(CB_Data$rideable_type, 
#                    CB_Data$rideable_type == "electric_bike",
#                    "Electric Bike")
head(CB_Data)
```
Now, calculate actual duration of ride (in minutes) as well as year and month.
```{r}
setDT(CB_Data)[, year := format(as.Date(started_at), "%Y") ]
setDT(CB_Data)[, month := format(as.Date(started_at), "%m") ]
CB_Data$duration = as.numeric(
                              difftime(
                                CB_Data$ended_at, 
                                CB_Data$started_at, 
                                units ="mins"
                                )
                              )
head(CB_Data)
```
Figure out the distances being traveled. Since calculating actual roadmap distance through Google Maps or Bing require use of APIs that cost money and place limitations on daily computations, that process shall be forgone in favor of "as the crow flies", which will provide a rough estimate of travel distance in order to figure out speed.

Convert meters to miles.
```{r}
# https://stackoverflow.com/questions/49532911/calculate-distance-longitude-latitude-of-multiple-in-dataframe-r
CB_Data$distance <- 0.000621371 * distHaversine(
                              cbind(CB_Data$start_lng, CB_Data$start_lat), 
                              cbind(CB_Data$end_lng, CB_Data$end_lat) 
                              )
head(CB_Data)
```
Find the zip codes associated with the coordinate data. This will help with ascertaining neighborhood and borough locations for simplicity.
```{r}
CB_Data$start_zip <- revgeo(CB_Data$start_lng, CB_Data$start_lat,
                            provider = NULL, API = NULL, output = "hash",
                            item = "zip")
CB_Data$end_zip <- revgeo(CB_Data$end_lng, CB_Data$end_lat,
                            provider = NULL, API = NULL, output = "hash",
                            item = "zip")
# start_zip <- revgeo(-73.95772, 40.77863,
#                             provider = NULL, API = NULL, output = "hash",
#                             item = "zip")
# end_zip <- revgeo(-73.95578, 40.77280,
#                             provider = NULL, API = NULL, output = "hash",
#                             item = "zip")
# start_zip
# end_zip

head(CB_Data)
```

